ACNP 2024 MS ANXIETY

### 2024-08-29 scripts/MSAnxiety_manuscript_audrey_replication_prop.Rmd
changed vol to prop - more approp for uf

#define group and make column for people who do have uncinate lesions and those who don't
df_demo_and_fascicles_no_unclass_anxiety_dose <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(anxiety_dose != -1) %>% #n=371
  mutate(has.uncinate.lesion = ifelse(mean_UF_vol >0, 1, 0)) #anywhere where there is *some* overlap counts as a 1 - they have an uncinate lesion, otherwise 0

@@

#### Uncinate by anxiety diagnosis 
```{r uncinate_by_dx}
### Having found that there is no association between the number of uncinate lesions and the anxiety dose, I next explored whether the *burden* of uncinate lesions differs by diagnosis/group

@@

### Results:  
#define group
  uncinate_by_dx_df <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "Has.Anxietydx.OR.Antianxiety.Meds") %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  mutate(Diagnosis = ifelse(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed=="true_healthy", "MS-Anxiety", "MS+Anxiety")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
 # dplyr::select(subject, osex, Diagnosis, PAT_AGE_AT_EXAM, UF_R_vol, UF_L_vol, F_R_vol, F_L_vol, mean_UF_vol) %>%
  ungroup() #n=122

#bilateral UF ~ Diagnosis: DiagnosisMS+Anxiety   816.97     404.08   2.022   0.0454 * 
disease_mean_UF_vol_gam <- gam(mean_UF_vol ~ Diagnosis + osex+s(PAT_AGE_AT_EXAM, k = 4, fx = F), data = uncinate_by_dx_df)
summary(disease_mean_UF_vol_gam)

#print summary stats
 gam_summary_stats(disease_mean_UF_vol_gam)
```

#### Uncinate by Anxiety Dose 
```{r uncinateByAnxietyDose}

@@

  #bilat anxiety dose + sex + age: anxiety_dose    383.3      184.3   2.080   0.0383 *  
  anxietyGrouping_mean_UF_by_sex_gam <- gam(mean_UF_vol ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = F), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_mean_UF_by_sex_gam)

  #print summary stats
 gam_summary_stats(anxietyGrouping_mean_UF_by_sex_gam)

@@

###### Laterality in uncinate
```{r laterality}
  #laterality 
  #overall - t = -8.2035, df = 661.94, p-value = 1.217e-15
  right_vs_left_uncinate <- t.test(df_demo_and_fascicles_no_unclass_anxiety_dose$UF_L_vol, df_demo_and_fascicles_no_unclass_anxiety_dose$UF_R_vol)
  print(right_vs_left_uncinate)

  #healthy - t = -4.3999, df = 171.31, p-value = 1.898e-05
   right_vs_left_uncinate_healthy <- t.test(df_demo_and_fascicles_no_unclass_anxiety_dose$UF_L_vol[df_demo_and_fascicles_no_unclass_anxiety_dose$anxiety_dose == 0], df_demo_and_fascicles_no_unclass_anxiety_dose$UF_R_vol[df_demo_and_fascicles_no_unclass_anxiety_dose$anxiety_dose == 0])
  print(right_vs_left_uncinate_healthy)

  #anxietydx or meds, t = -6.5586, df = 443.4, p-value = 1.516e-10
  right_vs_left_uncinate_anxiety_or_meds <- t.test(df_demo_and_fascicles_no_unclass_anxiety_dose$UF_L_vol[df_demo_and_fascicles_no_unclass_anxiety_dose$anxiety_dose == 1], df_demo_and_fascicles_no_unclass_anxiety_dose$UF_R_vol[df_demo_and_fascicles_no_unclass_anxiety_dose$anxiety_dose == 1])
  print(right_vs_left_uncinate_anxiety_or_meds)

  #anxious+meds,t = -2.261, df = 42.723, p-value = 0.02891
   right_vs_left_uncinate_anxiety_and_meds <- t.test(df_demo_and_fascicles_no_unclass_anxiety_dose$UF_L_vol[df_demo_and_fascicles_no_unclass_anxiety_dose$anxiety_dose == 2], df_demo_and_fascicles_no_unclass_anxiety_dose$UF_R_vol[df_demo_and_fascicles_no_unclass_anxiety_dose$anxiety_dose == 2])
  print(right_vs_left_uncinate_anxiety_and_meds)
```

#### Sensitivity
###### UF by depGroupVar (N.S)
```{r UF_by_depGroupVar}
# Specificity - want to make sure the UF findings are specific to anxiety, so check to see if same findings exist in depression
### Reviewed previous pull from BP paper, finding was also consistent there -> no diff in uncinate in depressed vs non depressed

@@

#mean UF, depDiagnosisMS+noDepression   -386.4      232.8  -1.660    0.098 .
dep_mean_UF_gam <- gam(mean_UF_vol ~ depDiagnosis + osex + s(PAT_AGE_AT_EXAM, k=4, fx=F), data = df_dep)
summary(dep_mean_UF_gam)

  #print summary stats
 gam_summary_stats(dep_mean_UF_gam)
```

###### Fornix by Anxiety Diagnosis and Dose (N.S)
```{r fornixByAnxietyDose}

@@

##### healthy vs sick
#Fornix gam Diagnosis; DiagnosisMS+Anxiety    283.0      375.6   0.753   0.4528   - NS
anxietyDiagnosis_mean_F_by_sex_gam <- gam(mean_fornix_volume ~ Diagnosis + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data = df_empi_dx_network_proportions)
summary(anxietyDiagnosis_mean_F_by_sex_gam)

 #print summary stats
 gam_summary_stats(anxietyDiagnosis_mean_F_by_sex_gam)

@@

  #gam, Fornix anxiety_dose    173.1      166.7   1.038    0.300   - NS
  anxietyGrouping_mean_F_by_sex_gam <- gam(mean_fornix_volume ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_mean_F_by_sex_gam)

 #print summary stats
 gam_summary_stats(anxietyGrouping_mean_F_by_sex_gam)
```

@@

## Make df for Anxiety Diagnosis
#define group
  uncinate_by_dx_df <- df_demo_and_fascicles_no_steroids_or_ifn %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "Has.Anxietydx.OR.Antianxiety.Meds") %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  mutate(Diagnosis = ifelse(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed=="true_healthy", "MS-Anxiety", "MS+Anxiety")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
 # dplyr::select(subject, osex, Diagnosis, PAT_AGE_AT_EXAM, UF_R_vol, UF_L_vol, F_R_vol, F_L_vol, mean_UF_vol) %>%
  ungroup()

#bilateral UF ~ Diagnosis: DiagnosisMS+Anxiety    718.8      413.2    1.74   0.0846 .   trend, likely due to sample size
disease_mean_UF_vol_gam <- gam(mean_UF_vol ~ Diagnosis + osex+s(PAT_AGE_AT_EXAM, k = 4, fx = F), data = uncinate_by_dx_df)
summary(disease_mean_UF_vol_gam)

 #print summary stats
 gam_summary_stats(disease_mean_UF_vol_gam)

@@

  #bilat anxiety dose + sex + age: anxiety_dose: anxiety_dose    358.3      188.3   1.903   0.0578 .   (trend, likely due to sample size)   
  anxietyGrouping_mean_UF_by_sex_gam <- gam(mean_UF_vol ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = F), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_mean_UF_by_sex_gam)

   #print summary stats
 gam_summary_stats(anxietyGrouping_mean_UF_by_sex_gam)
```

@@

# mean uncinate L - osex.L          -152.317    174.353  -0.874    0.383 ns
disease_between_sexes_UF_mean <- lm(mean_UF_vol ~ osex+ PAT_AGE_AT_EXAM, data = df_demo_and_fascicles_no_unclass_anxiety_dose)
summary(disease_between_sexes_UF_mean)

### Age*Disease

#### Anxiety diagnosis
```{r anxiety_dx_x_age}

#purpose of this analysis is to evaluate the relationship between disease (cross sectional) by diagnosis as well as by anxiety dose
#hypothesize that anxiety vs nonanxiety will have worse disease overall, and that this may be related to anxiety dose
#start by defining new variables and a new data frame that contains anxiety dose (as factor and as ordinal), as well as diagnosis (no anxiety vs severe anxiety)

#define group
df_demo_and_fascicles_no_unclass_anxiety_dose <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(anxiety_dose != -1) %>%
  mutate(anxiety_dose_factor = ordered(anxiety_dose)) %>%
  mutate(Diagnosis = oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  mutate(Diagnosis = ifelse(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed=="true_healthy", "MS-Anxiety", "MS+Anxiety")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) #n=371

### diagnosis by age interaction?
#overall - DiagnosisMS+Anxiety  0.03966    0.02329   1.703   0.0893 .  NS for diagnosis
diagnosis_by_age <- lm(proportion_volume_lost_per_total_network_size_sum ~ Diagnosis*PAT_AGE_AT_EXAM, data = df_demo_and_fascicles_no_unclass_anxiety_dose)
summary(diagnosis_by_age)

#gam diagnosis - SPLINE SIGNIFICANT
#s(PAT_AGE_AT_EXAM):DiagnosisMS-Anxiety 1.581  1.930 5.026 0.01362 * 
#s(PAT_AGE_AT_EXAM):DiagnosisMS+Anxiety 1.994  2.385 5.249 0.00313 **
diagnosis_by_age_gam <- gam(proportion_volume_lost_per_total_network_size_sum ~ Diagnosis+s(PAT_AGE_AT_EXAM, by = Diagnosis, k=4, fx = F), data = df_demo_and_fascicles_no_unclass_anxiety_dose, method = "REML")
summary(diagnosis_by_age_gam)
summary(diagnosis_by_age_gam)$s.table

  #print summary stats
 gam_summary_stats(diagnosis_by_age_gam)

#plot proportion_volume_lost_per_total_network_size_sum
plot_smooths_proportion_volume_lost_per_total_network_size_sum <- plot_smooths(
  model = diagnosis_by_age_gam,
  series = PAT_AGE_AT_EXAM,
  comparison = Diagnosis
) +
  theme(legend.position = "top")
plot_smooths_proportion_volume_lost_per_total_network_size_sum

#uncinate mean, DiagnosisMS+Anxiety    420.3      224.9   1.869   0.0624 . NS
#s(PAT_AGE_AT_EXAM):DiagnosisMS-Anxiety 1.007  1.014 23.98 2.07e-06 ***
#s(PAT_AGE_AT_EXAM):DiagnosisMS+Anxiety 1.000  1.000 17.15 4.36e-05 ***

diagnosis_by_age_mean_UF_gam <- gam(mean_UF_vol ~ Diagnosis+s(PAT_AGE_AT_EXAM, by = Diagnosis, k=4, fx = F), data = df_demo_and_fascicles_no_unclass_anxiety_dose, method = "REML")
summary(diagnosis_by_age_mean_UF_gam)
summary(diagnosis_by_age_mean_UF_gam)$s.table
#plot(diagnosis_by_age_UF_L_gam)

  #print summary stats
 gam_summary_stats(diagnosis_by_age_mean_UF_gam)

#gam UF mean plot

plot_smooths_anxiety_diagnosis_by_age_mean_UF_vol <- plot_smooths(
  model = diagnosis_by_age_mean_UF_gam,
  series = PAT_AGE_AT_EXAM,
  comparison = Diagnosis
) +
  theme(legend.position = "top")
plot_smooths_anxiety_diagnosis_by_age_mean_UF_vol

## Sex diff trajectories by age
#overall - osex.L                  0.066402   0.075006   0.885  0.37658  NS
sex_by_age <- lm(proportion_volume_lost_per_total_network_size_sum ~ osex*PAT_AGE_AT_EXAM, data = df_demo_and_fascicles_no_unclass_anxiety_dose)
summary(sex_by_age)

#### Anxiety DOSE 
```{r anxietydosexage}

#purpose of this analysis is to evaluate the relationship between disease (cross sectional) by diagnosis as well as by anxiety dose
#hypothesize that anxiety vs nonanxiety will have worse disease overall, and that this may be related to anxiety dose
#start by defining new variables and a new data frame that contains anxiety dose (as factor and as ordinal), as well as diagnosis (no anxiety vs severe anxiety)

#define group
df_demo_and_fascicles_no_unclass_anxiety_dose <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(anxiety_dose != -1) %>%
  mutate(anxiety_dose_factor = ordered(anxiety_dose)) %>%
  mutate(Diagnosis = oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  mutate(Diagnosis = ifelse(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed=="true_healthy", "MS-Anxiety", "MS+Anxiety")) %>%
  mutate(Diagnosis = as.factor(Diagnosis))

### diagnosis by age interaction?
#overall - NS
diagnosis_by_age <- lm(proportion_volume_lost_per_total_network_size_sum ~ anxiety_dose*PAT_AGE_AT_EXAM, data = df_demo_and_fascicles_no_unclass_anxiety_dose)
summary(diagnosis_by_age)

#gam dose; s(PAT_AGE_AT_EXAM):anxiety_dose 2.058  2.452 4.972 0.00315 **
diagnosis_by_age_gam <- gam(proportion_volume_lost_per_total_network_size_sum ~ anxiety_dose+s(PAT_AGE_AT_EXAM, by = anxiety_dose, k=4, fx = F), data = df_demo_and_fascicles_no_unclass_anxiety_dose, method = "REML")
summary(diagnosis_by_age_gam)
summary(diagnosis_by_age_gam)$s.table
#plot(diagnosis_by_age_gam)

  #print summary stats
 gam_summary_stats(diagnosis_by_age_gam)

#gam factor - there is a linear dose effect per network size sum, as well as a sig age by anxiety relationship, but only in sick people
#anxiety_dose_factor.L  0.07788    0.03209   2.427   0.0157 * 
#s(PAT_AGE_AT_EXAM):anxiety_dose_factor1 1.459  1.765 5.736  0.0138 *
#s(PAT_AGE_AT_EXAM):anxiety_dose_factor2 1.972  2.372 2.128  0.1123  
diagnosis_by_age_gam_factor <- gam(proportion_volume_lost_per_total_network_size_sum ~ anxiety_dose_factor+s(PAT_AGE_AT_EXAM, by = anxiety_dose_factor, k=4, fx = F), data = df_demo_and_fascicles_no_unclass_anxiety_dose, method = "REML")
summary(diagnosis_by_age_gam_factor)
summary(diagnosis_by_age_gam_factor)$s.table
#plot(diagnosis_by_age_gam_factor)

  #print summary stats
 gam_summary_stats(diagnosis_by_age_gam_factor)

#plot proportion_volume_lost_per_total_network_size_sum
plot_smooths_proportion_volume_lost_per_total_network_size_sum <- plot_smooths(
  model = diagnosis_by_age_gam_factor,
  series = PAT_AGE_AT_EXAM,
  comparison = anxiety_dose_factor
) +
  theme(legend.position = "top")
plot_smooths_proportion_volume_lost_per_total_network_size_sum

#uncinate mean
#dose
#s(PAT_AGE_AT_EXAM):anxiety_dose 1.003  1.005 12.59 0.000427 ***
diagnosis_by_age_mean_UF_gam <- gam(mean_UF_vol ~ anxiety_dose+s(PAT_AGE_AT_EXAM, by = anxiety_dose, k=4, fx = F), data = df_demo_and_fascicles_no_unclass_anxiety_dose, method = "REML")
summary(diagnosis_by_age_mean_UF_gam)
summary(diagnosis_by_age_mean_UF_gam)$s.table

#factor
#s(PAT_AGE_AT_EXAM):anxiety_dose_factor1 1.002  1.004 17.462 3.69e-05 ***
#s(PAT_AGE_AT_EXAM):anxiety_dose_factor2 1.002  1.005  0.029    0.872 
diagnosis_by_age_mean_UF_gam_factor <- gam(mean_UF_vol ~ anxiety_dose_factor+s(PAT_AGE_AT_EXAM, by = anxiety_dose_factor, k=4, fx = F), data = df_demo_and_fascicles_no_unclass_anxiety_dose, method = "REML")
summary(diagnosis_by_age_mean_UF_gam_factor)
summary(diagnosis_by_age_mean_UF_gam_factor)$s.table
#plot(diagnosis_by_age_UF_L_gam_factor)

  #print summary stats
 gam_summary_stats(diagnosis_by_age_mean_UF_gam_factor)

#gam mean

plot_smooths_anxiety_dose_by_age_mean_UF_factor <- plot_smooths(
  model = diagnosis_by_age_mean_UF_gam_factor,
  series = PAT_AGE_AT_EXAM,
  comparison = anxiety_dose_factor
) +
  theme(legend.position = "top")
plot_smooths_anxiety_dose_by_age_mean_UF_factor

## Sex diff trajectories by age
#overall -osex.L                  0.066402   0.075006   0.885  0.37658  NS
sex_by_age <- lm(proportion_volume_lost_per_total_network_size_sum ~ osex*PAT_AGE_AT_EXAM, data = df_demo_and_fascicles_no_unclass_anxiety_dose)
summary(sex_by_age)

### Individual Fascicle disease

#### healthy vs super sick - values that fdr correct, most not anxiety net
```{r individual_fascicle_disease_healthy_vs_super_sick}

df_demo_and_fascicles_no_unclass_anxiety_dose <- df_demo_and_fascicles %>%
  mutate(anxiousAndMeds_or_healthy = oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  filter(anxiousAndMeds_or_healthy %in% c("true_healthy", "Has.Anxiety.And.On.Anxiety.Meds.Inc.Dep")) %>%
  mutate(anxiousAndMeds_or_healthy = factor(anxiousAndMeds_or_healthy, levels=c("true_healthy", "Has.Anxiety.And.On.Anxiety.Meds.Inc.Dep")))

#lm
fascicle_lm <- lapply(fascicle_names_w_vol_suffix, function(x) 
{
  lm(substitute(i ~ anxiousAndMeds_or_healthy + PAT_AGE_AT_EXAM, list(i = as.name(x))), data = df_demo_and_fascicles_no_unclass_anxiety_dose)
})
names(fascicle_lm) <- fascicle_names_w_vol_suffix

#anova
fascicle_lm_anova <- lapply(fascicle_lm, anova)

#uncorrected, n=33
fascicle_anova_lm_p <- sapply(fascicle_lm_anova, function(v) v$"Pr(>F)"[1])
fascicle_anova_lm_p05_unc <- as.data.frame(fascicle_anova_lm_p[fascicle_anova_lm_p < 0.05])
print(fascicle_anova_lm_p05_unc)

#fdr corrected, n = 11
fascicle_anova_lm_fdr <- fdr_anova_generic(fascicle_lm_anova, 1, print_unc_df = TRUE)
print(fascicle_anova_lm_fdr)

#gam
fascicle_gam <- lapply(fascicle_names_w_vol_suffix, function(x) 
{
  gam(substitute(i ~ anxiousAndMeds_or_healthy + +osex+ s(PAT_AGE_AT_EXAM, k=4, fx=F), list(i = as.name(x))), data = df_demo_and_fascicles_no_unclass_anxiety_dose)
})
names(fascicle_gam) <- fascicle_names_w_vol_suffix

#anova
fascicle_anova_gam <- lapply(fascicle_gam, anova)

#uncorrected (38 items)
fascicle_anova_gam_p <- sapply(fascicle_anova_gam, function(v) v$p.pv[2])
print(as.data.frame(fascicle_anova_gam_p))
fascicle_anova_gam_p05_unc <- as.data.frame(fascicle_anova_gam_p[fascicle_anova_gam_p < 0.05])
print(fascicle_anova_gam_p05_unc)

#fdr corrected (18 items correct)
fascicle_anova_gam_fdr <- round(p.adjust(fascicle_anova_gam_p, method = "fdr"),3)
fascicle_anova_gam_fdr_p05 <- as.data.frame(fascicle_anova_gam_fdr[fascicle_anova_gam_fdr < 0.05])
print(fascicle_anova_gam_fdr_p05) 
```

#### Parametric all fascicles
```{r individual_fascicle_disease_anxiety_dose}

### 
df_demo_and_fascicles_no_unclass_anxiety_dose <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(anxiety_dose != -1) %>%
  mutate(anxiety_dose_factor = ordered(anxiety_dose))

#lm
fascicle_lm <- lapply(fascicle_names_w_vol_suffix, function(x) 
{
  lm(substitute(i ~ anxiety_dose + PAT_AGE_AT_EXAM, list(i = as.name(x))), data = df_demo_and_fascicles_no_unclass_anxiety_dose)
})
names(fascicle_lm) <- fascicle_names_w_vol_suffix

#anova
fascicle_anova_lm <- lapply(fascicle_lm, anova)

#fdr corrected, n = 1
fascicle_anova_lm_fdr <- fdr_anova_generic(fascicle_anova_lm, 1, print_unc_df = TRUE)
print(fascicle_anova_lm_fdr)

#uncorrected, n = 19
fascicle_anova_lm_p <- sapply(fascicle_anova_lm, function(v) v$"Pr(>F)"[1])
fascicle_anova_lm_p05_unc <- as.data.frame(fascicle_anova_lm_p[fascicle_anova_lm_p < 0.05])
print(fascicle_anova_lm_p05_unc)

#gam
fascicle_gam <- lapply(fascicle_names_w_vol_suffix, function(x) 
{
  gam(substitute(i ~ anxiety_dose + osex+s(PAT_AGE_AT_EXAM, k=4, fx=F), list(i = as.name(x))), data = df_demo_and_fascicles_no_unclass_anxiety_dose)
})
names(fascicle_gam) <- fascicle_names_w_vol_suffix

#anova
fascicle_anova_gam <- lapply(fascicle_gam, anova)

#uncorrected, n = 30
fascicle_anova_gam_p <- sapply(fascicle_anova_gam, function(v) v$p.pv[2])
print(as.data.frame(fascicle_anova_gam_p))
fascicle_anova_gam_p05_unc <- as.data.frame(fascicle_anova_gam_p[fascicle_anova_gam_p < 0.05])
print(fascicle_anova_gam_p05_unc)

#fdr corrected, n = 11
fascicle_anova_gam_fdr <- round(p.adjust(fascicle_anova_gam_p, method = "fdr"),3)
fascicle_anova_gam_fdr_p05 <- as.data.frame(fascicle_anova_gam_fdr[fascicle_anova_gam_fdr < 0.05])
print(fascicle_anova_gam_fdr_p05) 
```
### Plots

##### Overall disease, within and outside anxiety net
```{r}
#n = 372

@@

#bilateral UF ~ Diagnosis: DiagnosisMS+Anxiety   -187.4      885.1  -0.212    0.835 N.S.
disease_mean_UF_vol_gam <- gam(mean_UF_vol ~ Diagnosis + osex+s(PAT_AGE_AT_EXAM, k = 4, fx = F), data = uncinate_by_dx_df)
summary(disease_mean_UF_vol_gam)

    #print summary stats
 gam_summary_stats(disease_mean_UF_vol_gam)

>>>>

#define group and make column for people who do have uncinate lesions and those who don't
df_demo_and_fascicles_no_unclass_anxiety_dose <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(anxiety_dose != -1) %>% #n=371
  mutate(has.uncinate.lesion = ifelse(mean_UF_prop >0, 1, 0)) #anywhere where there is *some* overlap counts as a 1 - they have an uncinate lesion, otherwise 0

@@

#### Uncinate by anxiety diagnosis (vol and prop)
```{r uncinate_by_dx}
### Having found that there is no association between the number of uncinate lesions and the anxiety dose, I next explored whether the *burden* of uncinate lesions differs by diagnosis/group

@@

### Results:  
#define group
  uncinate_by_dx_df <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "Has.Anxietydx.OR.Antianxiety.Meds") %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  mutate(Diagnosis = ifelse(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed=="true_healthy", "MS-Anxiety", "MS+Anxiety")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  ungroup() #n=123

#bilateral UF ~ Diagnosis: DiagnosisMS+Anxiety   0.10624    0.05255   2.022   0.0455  * 
disease_mean_UF_prop_gam <- gam(mean_UF_prop ~ Diagnosis + osex+s(PAT_AGE_AT_EXAM, k = 4, fx = F), data = uncinate_by_dx_df)
summary(disease_mean_UF_prop_gam)

 #print summary stats
  gam_summary_stats(disease_mean_UF_prop_gam)
```

#### Uncinate by Anxiety Dose (vol and prop)
```{r uncinateByAnxietyDose}

@@

    #bilat anxiety dose + sex + age: anxiety_dose    0.04989    0.02397   2.082   0.0381 * 
 anxietyGrouping_mean_UF_by_sex_gam_prop <- gam(mean_UF_prop ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = F), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_mean_UF_by_sex_gam_prop)

  #print summary stats
  gam_summary_stats(anxietyGrouping_mean_UF_by_sex_gam_prop)

@@

###### Laterality in uncinate (prop)
```{r laterality}
  #laterality 
  #overall - t = -7.9761, df = 668.34, p-value = 6.564e-15
  right_vs_left_uncinate <- t.test(df_demo_and_fascicles_no_unclass_anxiety_dose$UF_L_prop, df_demo_and_fascicles_no_unclass_anxiety_dose$UF_R_prop)
  print(right_vs_left_uncinate)

  #healthy - t = -4.2898, df = 173.04, p-value = 2.969e-05
   right_vs_left_uncinate_healthy <- t.test(df_demo_and_fascicles_no_unclass_anxiety_dose$UF_L_prop[df_demo_and_fascicles_no_unclass_anxiety_dose$anxiety_dose == 0], df_demo_and_fascicles_no_unclass_anxiety_dose$UF_R_prop[df_demo_and_fascicles_no_unclass_anxiety_dose$anxiety_dose == 0])
  print(right_vs_left_uncinate_healthy)

  #anxietydx or meds, t = -6.3724, df = 447.66, p-value = 4.629e-10
  right_vs_left_uncinate_anxiety_or_meds <- t.test(df_demo_and_fascicles_no_unclass_anxiety_dose$UF_L_prop[df_demo_and_fascicles_no_unclass_anxiety_dose$anxiety_dose == 1], df_demo_and_fascicles_no_unclass_anxiety_dose$UF_R_prop[df_demo_and_fascicles_no_unclass_anxiety_dose$anxiety_dose == 1])
  print(right_vs_left_uncinate_anxiety_or_meds)

  #anxious+meds,t = -2.19, df = 43.079, p-value = 0.03398
   right_vs_left_uncinate_anxiety_and_meds <- t.test(df_demo_and_fascicles_no_unclass_anxiety_dose$UF_L_prop[df_demo_and_fascicles_no_unclass_anxiety_dose$anxiety_dose == 2], df_demo_and_fascicles_no_unclass_anxiety_dose$UF_R_prop[df_demo_and_fascicles_no_unclass_anxiety_dose$anxiety_dose == 2])
  print(right_vs_left_uncinate_anxiety_and_meds)
```

#### Sensitivity
###### UF by depGroupVar (vol and prop, N.S)
```{r UF_by_depGroupVar}
# Specificity - want to make sure the UF findings are specific to anxiety, so check to see if same findings exist in depression
### Reviewed previous pull from BP paper, finding was also consistent there -> no diff in uncinate in depressed vs non depressed

@@

 #mean UF, depDiagnosisMS+noDepression   -0.05029    0.03028  -1.661   0.0978 . 
dep_mean_UF_gam_prop <- gam(mean_UF_prop ~ depDiagnosis + osex + s(PAT_AGE_AT_EXAM, k=4, fx=F), data = df_dep)
summary(dep_mean_UF_gam_prop)

  #print summary stats
 gam_summary_stats(dep_mean_UF_gam_prop)
```

###### Fornix by Anxiety Diagnosis and Dose (vol and prop, N.S)
```{r fornixByAnxietyDose}

@@

 ##### healthy vs sick - prop
#Fornix gam Diagnosis; DiagnosisMS+Anxiety    0.03690    0.05405   0.683   0.4961    - NS
anxietyDiagnosis_mean_F_by_sex_gam_prop <- gam(mean_fornix_prop ~ Diagnosis + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data = df_empi_dx_network_proportions)
summary(anxietyDiagnosis_mean_F_by_sex_gam_prop)

@@

  #gam, Fornix anxiety_dose   0.02303    0.02401   0.959    0.338    - NS
  anxietyGrouping_mean_F_by_sex_gam_prop <- gam(mean_fornix_prop ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_mean_F_by_sex_gam_prop)

 #print summary stats
 gam_summary_stats(anxietyDiagnosis_mean_F_by_sex_gam_prop)
 gam_summary_stats(anxietyGrouping_mean_F_by_sex_gam_prop)
```

@@

## Make df for Anxiety Diagnosis
#define group
  uncinate_by_dx_df <- df_demo_and_fascicles_no_steroids_or_ifn %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "Has.Anxietydx.OR.Antianxiety.Meds") %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  mutate(Diagnosis = ifelse(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed=="true_healthy", "MS-Anxiety", "MS+Anxiety")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  ungroup()

 #bilateral UF ~ Diagnosis: DiagnosisMS+Anxiety    0.09339    0.05373   1.738   0.0849 .    trend, likely due to sample size
disease_mean_UF_gam_prop <- gam(mean_UF_prop ~ Diagnosis + osex+s(PAT_AGE_AT_EXAM, k = 4, fx = F), data = uncinate_by_dx_df)
summary(disease_mean_UF_gam_prop)

@@

  #bilat anxiety dose + sex + age: anxiety_dose: anxiety_dose    0.04662    0.02448   1.904   0.0577 .    (trend, likely due to sample size)   
  anxietyGrouping_mean_UF_by_sex_gam_prop <- gam(mean_UF_prop ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = F), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_mean_UF_by_sex_gam_prop)

   #print summary stats
   gam_summary_stats(disease_mean_UF_gam_prop)
   gam_summary_stats(anxietyGrouping_mean_UF_by_sex_gam_prop)
```

@@

# mean uncinate L - osex.L          -0.019879   0.022675  -0.877    0.381  ns
disease_between_sexes_UF_mean_prop <- lm(mean_UF_prop ~ osex+ PAT_AGE_AT_EXAM, data = df_demo_and_fascicles_no_unclass_anxiety_dose)
summary(disease_between_sexes_UF_mean_prop)
```

### Plots

##### Overall disease in brain
```{r}
#n = 372

@@

#bilateral UF prop ~ Diagnosis: DiagnosisMS+Anxiety -0.02490    0.11510  -0.216   0.8317 
disease_mean_UF_prop_gam <- gam(mean_UF_prop ~ Diagnosis + osex+s(PAT_AGE_AT_EXAM, k = 4, fx = F), data = uncinate_by_dx_df)
summary(disease_mean_UF_prop_gam)

    #print summary stats
 gam_summary_stats(disease_mean_UF_prop_gam)
